MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-06-04 MAESTROeX_I/intro.tex
add some more refs

Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, terrestrial atmospheric modeling \citep{duarte2015low},
and elastic solids \citep{abbate2017all}.
For astrophysical applications, a number of approaches to modeling low Mach number flows have been developed in recent years.
These include semi-implicit all-Mach number solvers, where the Euler equations are split into an acoustic part and an advective part \citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
The fast acoustic waves are then solved using implicit time integration, while the slow material waves are solved explicitly.
Another approach is to use preconditioned all-Mach number solvers \citep{Miczek2014,Barsukow2016}, where the numerical flux is multiplied by a preconditioning matrix.
This reduces the stiffness of the system at low Mach numbers, while retaining the correct scaling behavior. In the reduced speed of sound technique (RSST) and related methods, the speed of sound is artificially reduced by including a suitable scaling factor in the continuity equation, reducing the restriction on the size of the timestep \citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
The MUSIC code uses fully implicit time integration for the compressible Euler equations, which therefore allows for arbitrarily large timesteps \citep{Viallet2011,Viallet2015,Goffrey2016}.

Previously, we developed the low Mach number astrophysical solver, MAESTRO.
MAESTRO is a block-structured, Cartesian grid finite-volume, adaptive mesh refinement (AMR) code that has been successfully used for many years for a number of applications, detailed below.
Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited to, astrophysical settings with significant atmospheric stratification.
This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
MAESTRO uses an explicit Godunov approach for advection, a stiff ODE package (VODE) \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
The time-advancement strategy uses Strang splitting to couple the thermodynamic variables, and a second-order projection method to integrate the velocity equations subject to a divergence constraint.
Central to the algorithm is a time-varying, one-dimensional stratified background (or base) state density and pressure held in hydrostatic equilibrium that couple to the full Cartesian grid solution.
The original MAESTRO code was developed in the pure-Fortran90 FBoxLib software framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.

@@

Since then, there have been many scientific investigations using MAESTRO, which have included additional algorithmic enhancements.  Topics include:
\begin{itemize}
\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
\item Convection in massive stars \citep{Gilet:2013}.
\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
%  In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
%  In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
\end{itemize}

>>>>

Low Mach number models have been developed for a variety of contexts
including combustion \citep{day2000numerical}, terrestrial atmospheric
modeling \citep{durran:1989,oneill:2014,duarte2015low}, and elastic
solids \citep{abbate2017all}.  For astrophysical applications, a
number of approaches to modeling low Mach number flows have been
developed in recent years.  These include methods atmospheric model similar to
what is presented here~\cite{Lin:2006}.  There are also  semi-implicit all-Mach
number solvers, where the Euler equations are split into an acoustic
part and an advective part
\citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
The fast acoustic waves are then solved using implicit time
integration, while the slow material waves are solved explicitly.
Another approach is to use preconditioned all-Mach number solvers
\citep{Miczek2014,Barsukow2016}, where the numerical flux is
multiplied by a preconditioning matrix.  This reduces the stiffness of
the system at low Mach numbers, while retaining the correct scaling
behavior. In the reduced speed of sound technique (RSST) and related
methods, the speed of sound is artificially reduced by including a
suitable scaling factor in the continuity equation, reducing the
restriction on the size of the timestep
\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
Finally, there are fully implicit time integration codes for the
compressible Euler equations
\citep{Viallet2011,kifonidis:2012,Viallet2015,Goffrey2016}.  The MUSIC code
uses fully implicit time integration for the compressible Euler
equations, which therefore allows for arbitrarily large timesteps.

Previously, we developed the low Mach number astrophysical solver, MAESTRO.
MAESTRO is a block-structured, Cartesian grid finite-volume, adaptive mesh refinement (AMR) code that has been successfully used for many years for a number of applications, detailed below.
Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited to, astrophysical settings with significant atmospheric stratification.
This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
MAESTRO uses an explicit Godunov approach for advection, a stiff ODE package (VODE) \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
The time-advancement strategy uses Strang splitting to couple the thermodynamic variables, and a second-order projection method to integrate the velocity equations subject to a divergence constraint.
Central to the algorithm is a time-varying, one-dimensional stratified background (or base) state density and pressure held in hydrostatic equilibrium that couple to the full Cartesian grid solution.
The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework (available in the same github organization as AMReX).

@@

Since then, there have been many scientific investigations using MAESTRO, which have included additional algorithmic enhancements.  Topics include:
\begin{itemize}
\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
\item Convection in massive stars \citep{Gilet:2013,gilkis:2016}.
\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
%  In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
%  In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
\end{itemize}
